Learning and comparing functional connectomes between populations

### 2012-12-24 paper.tex
Iter on paper

\caption{
Correlation matrices of rest time-series extracted from the 39 main
regions of the Varoquaux 2011 \cite{varoquaux2011} parcellation -- 
Upper left: regressing out CompCorr signals, as well as white matter and
CSF average signals and movement parameters. The insert shows the
connections restricted to the default mode network.
% XXX: should I outline the default mode network on the main matrix?
-- Upper right: regressing out only movement parameters. -- Lower right:
regressing out movement parameters and global signal mean.
% XXX: should this remark go here?
No frequency filtering was applied here.
\label{fig:correlation_matrices}
}
\end{figure}
%---------------------------------------------------------------------------}}}

@@

Finally, it is important to keep in mind that using resting-state to
study on-going activity is fragile to confounds such as movement
\cite{vandijk2012,power2011} or scanner noise, probably more so than
activation studies as it is harder to control the specificity of the
signal extracted. Special care must be taken in preprocessing strategies
\cite{vandijk2010,satterthwaite2012}.
%
In addition, improved specificity to BOLD signal can be enforced by
summarizing the signal in an ROI by a mean of the different voxels
weighted by the subject-specific grey matter probabilistic segmentation,
as output by \emph{e.g.} SPM's segmentation tool \cite{ashburner2005}, .

@@

\paragraph{Correlation and partial correlations}

Correlation, shrinkage of correlation \cite{ledoit2004,varoquaux2012},

>>>>

\caption{
Correlation matrices of rest time-series extracted from the 39 main
regions of the Varoquaux 2011 \cite{varoquaux2011} parcellation -- 
Upper left: regressing out CompCor signals, as well as white matter and
CSF average signals and movement parameters. The insert shows the
connections restricted to the default mode network.
% XXX: should I outline the default mode network on the main matrix?
-- Upper right: regressing out only movement parameters. -- Lower right:
regressing out movement parameters and global signal mean.
% XXX: should this remark go here?
No frequency filtering was applied here.
\label{fig:correlation_matrices}
}
\end{figure}
%---------------------------------------------------------------------------}}}

@@

Finally, it is important to keep in mind that using resting-state to
study on-going activity is fragile to confounds such as movement
\cite{vandijk2012,power2011} or scanner noise, and probably more so than
activation studies as it is harder to control the specificity of the
signal extracted. Special care must be taken in preprocessing strategies
\cite{vandijk2010,satterthwaite2012}.
%
In addition, improved specificity to BOLD signal can be enforced by
using only signal in voxels near gray-matter tissues. For this purpose,
we suggest summarizing the signal in an ROI by a mean of the different voxels
weighted by the subject-specific gray matter probabilistic segmentation,
as output by \emph{e.g.} SPM's segmentation tool \cite{ashburner2005}.

@@

\paragraph{Correlation and partial correlations}
%
Functional connectivity between the ROIs can be measured by computing the
correlation matrix of the extracted signals. An important and often
neglected point is that the sample correlation matrix, \emph{i.e.} the
correlation matrix obtained by plugging the observed signal in the
correlation matrix formula, is not the population correlation matrix,
\emph{i.e.} the correlation matrix of the data-generating process. If the
number of measurements was infinite, the two would coincide, however if
this number is not large compared to the number of connection (that
scales as the square of the number of ROIs), the sample correlation
matrix is a poor estimate of the underlying population correlation matrix.
In other words, the sample correlation matrix captures a lot of sampling
noise, intrinsic randomness that arises in the estimation of correlations
from short time series. Conclusions drawn from the sample correlation
matrix can easily reflect this estimation error.

Varoquaux \emph{et al.} \cite{varoquaux2010c} and Smith \emph{et al.}
\cite{smith2011} have shown respectively on rest fMRI and on realistic
simulations that a good choice of correlation matrix estimator could
recover the connectivity structure, where the sample correlation matrix
would fail. In general, the choice of a better estimate depends on the
settings and the end goals \cite{varoquaux2012,varoquaux2010b}, however
the Ledoit-Wolf shrinkage estimate \cite{ledoit2004} is a simple,
computationally-efficient, and parameter-less alternative that performs
uniformly better than the sample correlation matrix
\cite{varoquaux2012,varoquaux2010c} and should always be prefered.
